---
title: "FIG Analysis Update"
author: "Jesse Puka-Beals"
date: '`r format(Sys.Date(), "%m %d %Y")`'
output: 
  ioslides_presentation:
    incremental: true
    # logo: logo.png
---

```{css, echo=FALSE}
slides > slide {
  overflow-x: auto !important;
  overflow-y: auto !important;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, nlme, lsmeans, lme4, multcomp, multcompView,
               googlesheets4, googledrive, lubridate, reshape2)
```

```{r import, include=FALSE, cache=TRUE}

# #Copy the url of googlesheet and name that url as an object in R
driveurl<-"https://docs.google.com/spreadsheets/d/16qh8tlNDfsdnwmDvass7WY5EFLY_U0AzQZVDhPQMhjQ/edit#gid=0"
# 
# #read_sheet() is the function that reads in the data. 
dat<-read_sheet(driveurl)

```

```{r tidy, include=FALSE, cache=TRUE}
dat<-dat%>%
  mutate_if(is.character,as.factor) %>%
  mutate_at("year", as.factor) %>%
  mutate_at("age", as.integer) %>%
  mutate_at("block", as.factor) %>%
  mutate_at("plot", as.factor) %>%
  mutate_at("trt.code", as.factor)

dat$timing.1cut <- factor(dat$timing.1cut, levels = c("boot", "anthesis", "dough", "grain"))
dat$follow.cut <- factor(dat$follow.cut, levels = c("none", "september", "october"))
dat$site <- factor(dat$site, levels = c("I2", "R70", "R100"))
dat$treatment <- factor(dat$treatment, levels = c("BN", "BS", "BO","AN", "AS", "AO","DN", "DS", "DO","GN", "GS", "GO"))

#for total yield comparisons
dat$yield.total <- rowSums(dat[,c("yield.1cut","yield.2cut","yield.3cut")], na.rm=T)

#for forage quality comparisons 
#First cut
DM<-dat$dry.matter.1cut
CP<-dat$protein.1cut
NDF<-dat$NDF.1cut
NDFD<-dat$NDFD48.1cut
ADF<-dat$ADF.1cut

EE<-2.05
FA<-EE-1
Ash<-100-DM
NFC<-100-((0.93*NDF)+CP+EE+Ash)
NDFn<-NDF*0.93
NDFDp<-22.7+0.664*NDFD
TDN<-(NFC*.98)+(CP*.87)+(FA*.97*2.25)+(NDFn*NDFDp/100)-10
DMI<-(-2.318)+(.442*CP)-(.01*CP^2)-(.0638*TDN)+(.000922*TDN^2)+(.18*ADF)-(0.00196*ADF^2)-(0.00529*CP*ADF)

dat$RFQ.1cut<-DMI*TDN/1.23
dat$RFV.1cut<-DMI*((89.8-(0.779*ADF)))/1.29

#Second cut
DM<-dat$dry.matter.2cut
CP<-dat$protein.2cut
NDF<-dat$NDF.2cut
NDFD<-dat$NDFD48.2cut
ADF<-dat$ADF.2cut

EE<-2.05
FA<-EE-1
Ash<-100-DM
NFC<-100-((0.93*NDF)+CP+EE+Ash)
NDFn<-NDF*0.93
NDFDp<-22.7+0.664*NDFD
TDN<-(NFC*.98)+(CP*.87)+(FA*.97*2.25)+(NDFn*NDFDp/100)-10
DMI<-(-2.318)+(.442*CP)-(.01*CP^2)-(.0638*TDN)+(.000922*TDN^2)+(.18*ADF)-(0.00196*ADF^2)-(0.00529*CP*ADF)

dat$RFQ.2cut<-DMI*TDN/1.23
dat$RFV.2cut<-DMI*((89.8-(0.779*ADF)))/1.29

#Third cut
DM<-dat$dry.matter.3cut
CP<-dat$protein.3cut
NDF<-dat$NDF.3cut
NDFD<-dat$NDFD48.3cut
ADF<-dat$ADF.3cut

EE<-2.05
FA<-EE-1
Ash<-100-DM
NFC<-100-((0.93*NDF)+CP+EE+Ash)
NDFn<-NDF*0.93
NDFDp<-22.7+0.664*NDFD
TDN<-(NFC*.98)+(CP*.87)+(FA*.97*2.25)+(NDFn*NDFDp/100)-10
DMI<-(-2.318)+(.442*CP)-(.01*CP^2)-(.0638*TDN)+(.000922*TDN^2)+(.18*ADF)-(0.00196*ADF^2)-(0.00529*CP*ADF)

dat$RFQ.3cut<-DMI*TDN/1.23
dat$RFV.3cut<-DMI*((89.8-(0.779*ADF)))/1.29

#weighting the RFQ and RFV by their proportion of the final total forage for a plot
dat$RFQ.1cut.wt <- dat$yield.1cut/dat$yield.total*dat$RFQ.1cut 
dat$RFQ.2cut.wt <- dat$yield.2cut/dat$yield.total*dat$RFQ.2cut 
dat$RFQ.3cut.wt <- dat$yield.3cut/dat$yield.total*dat$RFQ.3cut 
dat$RFQ.total <- rowSums(dat[,c("RFQ.1cut.wt","RFQ.2cut.wt","RFQ.3cut.wt")], na.rm=T)

dat$RFV.1cut.wt <- dat$yield.1cut/dat$yield.total*dat$RFV.1cut 
dat$RFV.2cut.wt <- dat$yield.2cut/dat$yield.total*dat$RFV.2cut 
dat$RFV.3cut.wt <- dat$yield.3cut/dat$yield.total*dat$RFV.3cut
dat$RFV.total <- rowSums(dat[,c("RFV.1cut.wt","RFV.2cut.wt","RFV.3cut.wt")], na.rm=T)


#incorporating $$$ using Hunter et al.
#Value of the hay in $ per Mg
dat$value.1cut<-34+1.02*dat$RFV.1cut
dat$value.2cut<-34+1.02*dat$RFV.2cut
dat$value.3cut<-34+1.02*dat$RFV.3cut

#Economic net economic return in $ per ha
#converts from Mg to kg/ha (0.001)
#subtracts cost of harvest ($143.06 per ha)
dat$return.1cut<-dat$value.1cut*dat$yield.1cut*0.001-143.06
dat$return.2cut<-dat$value.2cut*dat$yield.2cut*0.001-143.06
dat$return.3cut<-dat$value.3cut*dat$yield.3cut*0.001-143.06
dat$return.total<-rowSums(dat[,c("return.1cut","return.2cut","return.3cut")], na.rm=T)

rm(DM,CP,NDF,NDFD,ADF,EE,FA,Ash,NFC,NDFn,NDFDp,TDN,DMI)
```


```{r QAQC, include=FALSE, cache=TRUE}
dat1 <- dat %>%
  filter(timing.1cut!="grain" & site!="I2")

dat1 <- dat1 %>%
  filter(id != 107)

options(scipen=999)

dat3 <- subset(dat1,env=="R70.2017" | env=="R100.2018" | env=="R100.2019")

```


```{r weather data import, include=FALSE, warning=FALSE, cache=TRUE}
weatherurl<-"https://docs.google.com/spreadsheets/d/1I-kHBXThb0VSwOgRF6ZwiyOtC9-BwCtsKVl7_Z7-XXs/edit#gid=0"

weather<-read_sheet(weatherurl)

weather1 <- weather %>%
  mutate_at("year", as.factor) %>%
  mutate(year = recode_factor(year,
                      `30` = "30 year average"))

```

## Narrative

* year 1: Plant IWG in fall

* year 2: IWG establishes, is harveseted once for grain

* year 3 + 4: IWG is treated as a forage

* *how should we manage the IWG in years 3 and 4?*

* *when should we harvest?*

* *how many times should we harvest?*

## Sites

 Site   Planted   Data collected      
------ --------- ----------------  
 *I2*   *2011*      *2017*
 R70    2015      2017-2018
 R100   2016      2018-2019

* Rosemount MN

## Data collection

* response variables
  + yield
  + quality
    - **$**
* treatments
  + timing.1cut
    - boot, anthesis, dough, *grain*
  + follow.cut
    - none, sept, sept+oct
* sample times
  + 1.cut, 2.cut, 3.cut, 
    - **.total**

## Site conditions 1

```{r}
ggplot(subset(weather, year!="30"), 
                   aes(x = as.Date(yday(date), "2017-01-01"), 
                     y = cum.precip, 
                     color = factor(year(date)))) +
  geom_path() +
  scale_x_date(date_breaks="months", date_labels="%b") +
  labs(x="",
       colour="",
       y="Cumulative precipitation \n(mm)") +
  theme_bw()

```

## Site conditions 2

```{r}
ggplot(subset(weather, year!="30"), 
                aes(x = as.Date(yday(date), "1970-01-01"), 
                     y = `Cumulative GDD`, 
                     color = factor(year(date)))) +
  geom_line() +
  scale_x_date(date_breaks="months", date_labels="%b") +
  labs(x="",
       colour="",
       y="Cumulative Growing\nDegree Days") +
  theme_bw()
```

## Timing year

```{r}
dat1.2017 <- subset(dat1, year=="2017")
dat1.2018 <- subset(dat1, year=="2018")
dat1.2019 <- subset(dat1, year=="2019")

cut1.dates.2017 <- yday(as.Date(unique(dat1.2017$date.1cut), "2017-01-01"))
cut1.dates.2018 <- yday(as.Date(unique(dat1.2018$date.1cut), "2017-01-01"))
cut1.dates.2019 <- yday(as.Date(unique(dat1.2019$date.1cut), "2017-01-01"))

ggplot(subset(weather, year!="30"), 
                   aes(x = as.Date(yday(date), "2017-01-01"), 
                     y = cum.precip, 
                     color = factor(year(date)))) +
  geom_path(alpha=0.5) +
  geom_path(aes(y=`Cumulative GDD`),
            linetype="dotted") +
  scale_x_date(date_breaks="months", date_labels="%b") +
  labs(x="",
       colour="",
       y="Cumulative precipitation/GDD") +
  theme_bw() +
  geom_vline(xintercept = as.Date(cut1.dates.2017, "2017-01-01"),
             color="red",
             linetype=2) +
  geom_vline(xintercept = as.Date(cut1.dates.2018, "2017-01-01"),
           color="green",
           linetype=2) +
    geom_vline(xintercept = as.Date(cut1.dates.2018, "2017-01-01"),
           color="blue",
           linetype=2) +
  geom_vline(xintercept=as.Date(c("2017-09-20", "2017-10-31")),
             linetype=2)

```
```{r}
ggplot(subset(weather, year!="30"), 
                   aes(x = as.Date(yday(date), "2017-01-01"), 
                     y = cum.precip, 
                     color = factor(year(date)))) +
  geom_path(alpha=0.5) +
  geom_path(aes(y=`Cumulative GDD`),
            linetype="dotted") +
  scale_x_date(date_breaks="months", date_labels="%b") +
  labs(x="",
       colour="",
       y="Cumulative precipitation/GDD",
       title=".1cut") +
  theme_bw() +
  geom_vline(xintercept = as.Date(cut1.dates.2017, "2017-01-01"),
             color="red",
             linetype=2) +
  geom_vline(xintercept = as.Date(cut1.dates.2018, "2017-01-01"),
           color="dark green",
           linetype=2) +
    geom_vline(xintercept = as.Date(cut1.dates.2019, "2017-01-01"),
           color="blue",
           linetype=2) +
  geom_vline(xintercept=as.Date(c("2017-09-20", "2017-10-31")),
             linetype=2) +
  scale_x_date(limits = as.Date(c(150, 225), "2017-01-01"))
```

## Diagnostics

* missing data
  - grain cuts
  - follow-up cuts @ R70.2018 & I2.2017
* problematic data
  - one outlier value at R100.2019 for dough cut
  - I2 

  
```{r, fig.height=2.5,fig.width=5, fig.align="center"}
lattice::bwplot(yield.1cut~timing.1cut|env,subset(dat,env=="R100.2019" & timing.1cut!="grain"))
```


## Analysis 1 {.smaller}

**null hypothesis testing**

* repeated measures
  - each plot is measured 1-3 times per year depending on assigned follow.cut treatment
  - each plot is measured over 2 consecutive years
* random effects
  - environment
  - site:block
* fixed effects
  - follow.cut
  - timing.1cut

## Analysis 2 {.smaller}

**regression**

* responses vs.
  - GDD accumulation
  - *precipitation accumulation?*
  - *hydrothermal time?*
  
<!-- ## Model fitting 1 -->

<!-- * Box plot response by environment -->
<!--     * identify if there's a consistent trend -->
<!--         * interaction drivers? -->
<!--     * estimate distribution and variance -->
<!-- * Distribution -->
<!--     * normal -->
<!--         - lmer -->
<!--     * gamma (poisson) -->
<!--         - glmer -->

<!-- ## Model fitting 2 -->

<!-- * Fit all effects into model -->
<!--     i) remove random effects not reducing residual variance -->
<!--     i) compare multiple models -->
<!--         * log liklihood -->
<!--         * AIC -->
<!--     i) null hypothesis test @ alpha=0.05 -->
<!--     i) post-hoc mean comparison -->
<!--         * tukey, z-statistic -->


## yield of first cut in second year of data collection

```{r, echo=TRUE}
car::Anova(lmer(yield.1cut~follow.cut*timing.1cut+
                  (1|site:block)+
                  (1|env),
           data=subset(dat1, field.year=="second")))
```
## quality of first cut in second year of data collection 

```{r, echo=TRUE}
car::Anova(lmer(RFQ.1cut~follow.cut*timing.1cut+
                  (1|site:block)+
                  (1|env),
           data=subset(dat1, field.year=="second")))
```


## quality of second cut in second year 

```{r, echo=TRUE}
car::Anova(lmer(RFQ.2cut~follow.cut*timing.1cut+
                  (1|block),
           data=subset(dat1, env=="R100.2019" & follow.cut!="none")))
```

## quality of second cut in both years {.smaller}

```{r, echo=TRUE}
car::Anova(lmer(RFQ.2cut~year*follow.cut+
                  (1|block),
           data=subset(dat1, site=="R100" & follow.cut!="none")))
```

- compare between years at same site since no obvious carryover effect
- conclude that no difference in quality in second cut between the first and second year at R100.

## the effect of multiple cuttings


  dataset                     yield.1cut    rfq.1cut    .2cut     .3cut
----------------------       ------------  ----------  --------  --------
 R70.2018+R100.2019                ns        ns        
 R100.2019                                               ns
 
Table:  Mixed effect model outputs

- Second year datasets do not find differences among follow.cut treatment levels due to the follow.cut treatments applied in the first year.

- no carryover effect from years observed


## follow.cut contribution to yield

```{r}

dat2 <- dat1
dat2$yield.1cut <- dat1$yield.1cut
dat2$yield.2cut <- dat1$yield.1cut+dat1$yield.2cut
dat2$yield.3cut <- dat1$yield.1cut+dat1$yield.2cut+dat1$yield.3cut
dat2.long <- reshape2::melt(data = dat2, id.vars = c("follow.cut"), measure.vars = c("yield.3cut", "yield.2cut","yield.1cut"))

dat2.long <- dat2.long %>%
  mutate(variable = recode(variable,
             yield.3cut = "October cut",
             yield.2cut = "September cut",
             yield.1cut = "Spring cut")) %>%
  mutate(follow.cut = recode(follow.cut,
             october = "september\n+\noctober"))


ggplot(data=dat2.long,
       aes(x=follow.cut, 
           y=value, 
           fill=variable, color=variable)) +
      geom_bar(stat="summary",
               fun="mean",
               position ="identity") +
  labs(y="Cumulative Mean Total Forage Yield \n(kg ha-1 yr-1)",
       x="Follow-Up Cuts",
       fill="",
       color="") +
  theme_classic() 

```

- NS differences

## follow.cut rfq

```{r}
dat1.long <- reshape2::melt(data = dat3, id.vars = c("follow.cut"), measure.vars = c("RFQ.3cut", "RFQ.2cut","RFQ.1cut"))

dat1.long <- dat1.long %>%
  mutate(variable = recode(variable,
             RFQ.1cut = "Spring cut",
             RFQ.2cut = "September cut",
             RFQ.3cut = "October cut")) %>%
  mutate(follow.cut = recode(follow.cut,
             october = "september\n+\noctober"))


ggplot(data=dat1.long,
       aes(x=follow.cut, 
           y=value, 
           fill=variable, color=variable)) +
      geom_bar(stat="summary",
               fun="mean",
               position ="dodge") +
  labs(y="Average Relative Forage Quality",
       x="Follow-Up Cuts",
       fill="",
       color="",
       title="Unweighted Forage Quality") +
  theme_classic()

```

```{r}
dat4 <- subset(dat3, follow.cut=="october")
dat4.long <- reshape2::melt(data = dat4, id.vars = c("follow.cut","site","block","env"), measure.vars = c("RFQ.1cut", "RFQ.2cut","RFQ.3cut"))

dat4.long <- dat4.long %>%
  mutate(variable = recode(variable,
             RFQ.1cut = "Spring cut",
             RFQ.2cut = "September cut",
             RFQ.3cut = "October cut")) %>%
  mutate(follow.cut = recode(follow.cut,
             october = "september\n+\noctober"))

dat4.long.tidy <- dat4.long %>%
  group_by(variable) %>%
  drop_na(value) %>%
  summarise(n=n(), mean=mean(value), 
            se=sd(value)/sqrt(n)) %>%
  as.data.frame() %>%
  mutate(tukey = c("b","a","b"))


ggplot(dat4.long.tidy, aes(x=variable, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.3,
                size=1) +
  geom_bar(stat = "identity", width=0.7) +
  geom_text(aes(label=tukey),
            vjust=-1.5,
            size=8) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 200)) + #adjust for y-axis range 
  theme_classic() +
  labs(y="Relative Forage Quality",
       x="",
       title="Subsetted data where follow.cut=sept+oct")
```


## follow.cut contribution to RFQ

```{r}

dat2 <- dat1
dat2$RFQ.1cut <- dat1$RFQ.1cut.wt
dat2$RFQ.2cut <- dat1$RFQ.2cut.wt
dat2$RFQ.3cut <- dat1$RFQ.3cut.wt
dat2.long <- reshape2::melt(data = dat2, id.vars = c("follow.cut"), measure.vars = c("RFQ.3cut", "RFQ.2cut","RFQ.1cut"))

dat2.long <- dat2.long %>%
  mutate(variable = recode(variable,
             RFQ.3cut = "October cut",
             RFQ.2cut = "September cut",
             RFQ.1cut = "Spring cut")) %>%
  mutate(follow.cut = recode(follow.cut,
             october = "september\n+\noctober"))


ggplot(data=dat2.long,
       aes(x=follow.cut, 
           y=value, 
           fill=variable, color=variable)) +
      geom_bar(stat="summary",
               fun="mean",
               position ="stack") +
  labs(y="Average Relative Forage Quality",
       x="Follow-Up Cuts",
       fill="",
       color="") +
  theme_classic()
```

- NS differences

## yield.1cut vs. timing.1cut

```{r}
yield.1cut.tidy <- dat1 %>%
  group_by(timing.1cut) %>%
  summarise(n=n(), mean=mean(yield.1cut), 
            se=sd(yield.1cut)/sqrt(n)) %>%
  as.data.frame() %>%
  mutate(tukey = c("b","a", "a"))

ggplot(yield.1cut.tidy, aes(x=timing.1cut, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.3,
                alpha=1,
                size=1) +
  geom_bar(stat = "identity", width=0.7) +
  geom_text(aes(label=tukey),
            vjust=-1.3,
            size=8) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 5000)) + #adjust for y-axis range 
  theme_classic() +
  labs(y="Forage Yield of the first cut of the Year \n(kg ha-1)",
       x="",
       title="Combined field years")

```

## yield.2cut vs. timing.1cut

```{r}
yield.2cut.tidy <- dat3 %>%
  group_by(timing.1cut) %>%
  drop_na(yield.2cut) %>%
  summarise(n=n(), mean=mean(yield.2cut), 
            se=sd(yield.2cut)/sqrt(n)) %>%
  as.data.frame() %>%
  mutate(tukey = c("a","b","c"))


ggplot(yield.2cut.tidy, aes(x=timing.1cut, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.3,
                alpha=1,
                size=1) +
  geom_bar(stat = "identity", width=0.7) +
  geom_text(aes(label=tukey),
            vjust=-1.5,
            size=8) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 2300)) + #adjust for y-axis range 
  theme_classic() +
  labs(y="Mean Forage Yield\nSecond Cut\n(kg ha-1 yr-1)",
       x="",
       title="Combined across three environments")
```

## yield.total vs. timing.1cut

```{r}
dat3 <- subset(dat1,env=="R70.2017" | env=="R100.2018" | env=="R100.2019")

yield.total.tidy <- dat3 %>%
  group_by(treatment) %>%
  summarise(n=n(), mean=mean(yield.total), 
            se=sd(yield.total)/sqrt(n)) %>%
  as.data.frame() %>%
  mutate(tukey = c("b","ab", "ab", "ab", "a", "a", "ab", "ab", "ab"))


ggplot(yield.total.tidy, aes(x=treatment, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.3,
                alpha=1,
                size=1) +
  geom_bar(stat = "identity", width=0.7) +
  geom_text(aes(label=tukey),
            vjust=-1.5,
            size=8) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 10000)) + #adjust for y-axis range 
  theme_classic() +
  labs(y="Forage Yield (kg ha-1)",
       x="",
       title="Combined across three environments")
```

## rfq.1cut vs timing.1cut

```{r}
RFQ.1cut.tidy <- dat1 %>%
  group_by(timing.1cut) %>%
  summarise(n=n(), mean=mean(RFQ.1cut), 
            se=sd(RFQ.1cut)/sqrt(n)) %>%
  as.data.frame() %>%
  mutate(tukey = c("a","b", "c"))

ggplot(RFQ.1cut.tidy, aes(x=timing.1cut, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.3,
                alpha=1,
                size=1) +
  geom_bar(stat = "identity", width=0.7) +
  geom_text(aes(label=tukey),
            vjust=-1.2,
            size=8) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 200)) + #adjust for y-axis range 
  theme_classic() +
  labs(y="Forage Yield (kg ha-1)",
       x="",
       title="")

```

## rfq.2cut vs. timing.1cut
```{r}
rfq.2cut.tidy <- dat3 %>%
  group_by(timing.1cut) %>%
  drop_na(RFQ.2cut) %>%
  summarise(n=n(), mean=mean(RFQ.2cut), 
            se=sd(RFQ.2cut)/sqrt(n)) %>%
  as.data.frame() %>%
  mutate(tukey = c("b","a","a"))


ggplot(rfq.2cut.tidy, aes(x=timing.1cut, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.3,
                alpha=1,
                size=1) +
  geom_bar(stat = "identity", width=0.7) +
  geom_text(aes(label=tukey),
            vjust=-1.5,
            size=8) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 200)) + #adjust for y-axis range 
  theme_classic() +
  labs(y="Relative Forage Quality\nof the Second Cut",
       x="",
       title="Combined across three environments")
```

- may want regression here rather than mean comparison

## rfq.3cut vs. timing.1cut

```{r, echo=TRUE}
car::Anova(glmer(RFQ.3cut~timing.1cut +
                  (1|env) + (1|site:block),
           data=dat3,
           family=Gamma(link="log")))
```


## rfq.total vs. treatment

```{r}

RFQ.total.tidy <- dat3 %>%
  group_by(treatment) %>%
  summarise(n=n(), mean=mean(RFQ.total), 
            se=sd(RFQ.total)/sqrt(n)) %>%
  as.data.frame() %>%
  mutate(tukey = c("a","a","a", "bc","b", "bc","c","bc","bc"))


ggplot(RFQ.total.tidy, aes(x=treatment, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.3,
                alpha=1,
                size=1) +
  geom_bar(stat = "identity", width=0.7) +
  geom_text(aes(label=tukey),
            vjust=-1.5,
            size=8) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 200)) + #adjust for y-axis range 
  theme_classic() +
  labs(y="Relative Forage Quality",
       x="",
       title="Combined across three environments")
```


## $ vs treatment


```{r}
nr.total.tidy <- dat3 %>%
  group_by(treatment) %>%
  summarise(n=n(), mean=mean(return.total), 
            se=sd(return.total)/sqrt(n)) %>%
  as.data.frame() %>%
  mutate(tukey = c("","","","","","","","",""))

ggplot(nr.total.tidy, aes(x=treatment, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.3,
                alpha=1,
                size=1) +
  geom_bar(stat = "identity", width=0.7) +
  geom_text(aes(label=tukey),
            vjust=-1.5,
            size=8) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 600)) + #adjust for y-axis range 
  theme_classic() +
  labs(y="Net returns \n($ ha-1)",
       x="",
       title="Combined across three environments",
       caption="Boot+Sept+Oct is only treatment that may not differ from zero")
```

- NS differences

## $ vs treatment | year {.smaller}

```{r}
nrt.lmer.yr <- lmer(return.total~treatment*year +
                  (1|site:block), #year random effect adds nothing
                dat3,
                REML=TRUE)
car::Anova(nrt.lmer.yr)
cld(emmeans(nrt.lmer.yr, ~treatment|year))
```

## regression {.smaller}

```{r, fig.height=3}
ggplot(dat1, aes(y=yield.1cut, x=gdd.1cut, color=year))+
  #facet_grid(~year)+
  geom_point()+
  geom_smooth(method="lm", se=F, show.legend=F, color="black")+
  geom_smooth(method="lm", se=F, show.legend=F, color="red", formula = y ~ x + I(x^2))
```

- We conclude that our maximum yield for the first forage cut is 4,103 kg ha dry forage biomass and that this is achievable when 1,468 GDD have accumulated, which appears to be after anthesis and before dough stage.

