---
title: "FIG Analysis Update"
author: "Jesse Puka-Beals"
date: '`r format(Sys.Date(), "%m %d %Y")`'
output: 
  ioslides_presentation:
    incremental: true
    logo: logo.png

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, nlme, lsmeans, lme4, multcomp, multcompView,
               googlesheets4, googledrive, lubridate, reshape2)
```

```{r import, include=FALSE, cache=TRUE}

# #Copy the url of googlesheet and name that url as an object in R
driveurl<-"https://docs.google.com/spreadsheets/d/16qh8tlNDfsdnwmDvass7WY5EFLY_U0AzQZVDhPQMhjQ/edit#gid=0"
# 
# #read_sheet() is the function that reads in the data. 
dat<-read_sheet(driveurl,
                gs4_deauth())

```

```{r tidy, include=FALSE, cache=TRUE}

dat<-dat%>%
  mutate_if(is.character,as.factor) %>%
  mutate_at("year", as.factor) %>%
  mutate_at("age", as.integer) %>%
  mutate_at("block", as.factor) %>%
  mutate_at("plot", as.factor) %>%
  mutate_at("trt.code", as.factor)

dat$timing.1cut <- factor(dat$timing.1cut, levels = c("boot", "anthesis", "dough", "grain"))
dat$follow.cut <- factor(dat$follow.cut, levels = c("none", "september", "october"))
dat$site <- factor(dat$site, levels = c("I2", "R70", "R100"))
dat$treatment <- factor(dat$treatment, levels = c("BN", "BS", "BO","AN", "AS", "AO","DN", "DS", "DO","GN", "GS", "GO"))

#for total yield comparisons
dat$yield.total <- rowSums(dat[,c("yield.1cut","yield.2cut","yield.3cut")], na.rm=T)

#for forage quality comparisons 
#First cut
DM<-dat$dry.matter.1cut
CP<-dat$protein.1cut
NDF<-dat$NDF.1cut
NDFD<-dat$NDFD48.1cut
ADF<-dat$ADF.1cut

EE<-2.05
FA<-EE-1
Ash<-100-DM
NFC<-100-((0.93*NDF)+CP+EE+Ash)
NDFn<-NDF*0.93
NDFDp<-22.7+0.664*NDFD
TDN<-(NFC*.98)+(CP*.87)+(FA*.97*2.25)+(NDFn*NDFDp/100)-10
DMI<-(-2.318)+(.442*CP)-(.01*CP^2)-(.0638*TDN)+(.000922*TDN^2)+(.18*ADF)-(0.00196*ADF^2)-(0.00529*CP*ADF)

dat$RFQ.1cut<-DMI*TDN/1.23
dat$RFV.1cut<-DMI*((89.8-(0.779*ADF)))/1.29

#Second cut
DM<-dat$dry.matter.2cut
CP<-dat$protein.2cut
NDF<-dat$NDF.2cut
NDFD<-dat$NDFD48.2cut
ADF<-dat$ADF.2cut

EE<-2.05
FA<-EE-1
Ash<-100-DM
NFC<-100-((0.93*NDF)+CP+EE+Ash)
NDFn<-NDF*0.93
NDFDp<-22.7+0.664*NDFD
TDN<-(NFC*.98)+(CP*.87)+(FA*.97*2.25)+(NDFn*NDFDp/100)-10
DMI<-(-2.318)+(.442*CP)-(.01*CP^2)-(.0638*TDN)+(.000922*TDN^2)+(.18*ADF)-(0.00196*ADF^2)-(0.00529*CP*ADF)

dat$RFQ.2cut<-DMI*TDN/1.23
dat$RFV.2cut<-DMI*((89.8-(0.779*ADF)))/1.29

#Third cut
DM<-dat$dry.matter.3cut
CP<-dat$protein.3cut
NDF<-dat$NDF.3cut
NDFD<-dat$NDFD48.3cut
ADF<-dat$ADF.3cut

EE<-2.05
FA<-EE-1
Ash<-100-DM
NFC<-100-((0.93*NDF)+CP+EE+Ash)
NDFn<-NDF*0.93
NDFDp<-22.7+0.664*NDFD
TDN<-(NFC*.98)+(CP*.87)+(FA*.97*2.25)+(NDFn*NDFDp/100)-10
DMI<-(-2.318)+(.442*CP)-(.01*CP^2)-(.0638*TDN)+(.000922*TDN^2)+(.18*ADF)-(0.00196*ADF^2)-(0.00529*CP*ADF)

dat$RFQ.3cut<-DMI*TDN/1.23
dat$RFV.3cut<-DMI*((89.8-(0.779*ADF)))/1.29

#weighting the RFQ and RFV by their proportion of the final total forage for a plot
dat$RFQ.1cut.wt <- dat$yield.1cut/dat$yield.total*dat$RFQ.1cut 
dat$RFQ.2cut.wt <- dat$yield.2cut/dat$yield.total*dat$RFQ.2cut 
dat$RFQ.3cut.wt <- dat$yield.3cut/dat$yield.total*dat$RFQ.3cut 
dat$RFQ.total <- rowSums(dat[,c("RFQ.1cut.wt","RFQ.2cut.wt","RFQ.3cut.wt")], na.rm=T)

dat$RFV.1cut.wt <- dat$yield.1cut/dat$yield.total*dat$RFV.1cut 
dat$RFV.2cut.wt <- dat$yield.2cut/dat$yield.total*dat$RFV.2cut 
dat$RFV.3cut.wt <- dat$yield.3cut/dat$yield.total*dat$RFV.3cut
dat$RFV.total <- rowSums(dat[,c("RFV.1cut.wt","RFV.2cut.wt","RFV.3cut.wt")], na.rm=T)

rm(DM,CP,NDF,NDFD,ADF,EE,FA,Ash,NFC,NDFn,NDFDp,TDN,DMI)
```

```{r QAQC, include=FALSE, cache=TRUE}
dat1 <- dat %>%
  filter(timing.1cut!="grain" & site!="I2")

dat1 <- dat1 %>%
  filter(id != 107)

options(scipen=999)
```

```{r weather data import, include=FALSE, warning=FALSE, cache=TRUE}
weatherurl<-"https://docs.google.com/spreadsheets/d/1I-kHBXThb0VSwOgRF6ZwiyOtC9-BwCtsKVl7_Z7-XXs/edit#gid=0"

weather<-read_sheet(weatherurl)

weather1 <- weather %>%
  mutate_at("year", as.factor) %>%
  mutate(year = recode_factor(year,
                      `30` = "30 year average"))

```

## Narrative

* year 1: Plant IWG in fall

* year 2: IWG establishes, is harveseted once for grain

* year 3 + 4: IWG is treated as a forage

* *how should we manage the IWG in years 3 and 4?*

* *when should we harvest?*

* *how many times should we harvest?*

## Design

 Site   Planted   Data collected      
------ --------- ----------------  
 I2     2011      2017
 R70    2015      2017-2018
 R100   2016      2018-2019

* all sites in Rosemount MN

## Data collection

* response variables
  + yield
  + quality
    - **$**
* treatments
  + timing.1cut
    - boot, dough, anthesis, *grain*
  + follow.cut
    - none, sept, sept+oct
* sample times
  + 1.cut, 2.cut, 3.cut, 
    - **.total**

## Site conditions 1

```{r}
ggplot(subset(weather, year!="30"), 
                   aes(x = as.Date(yday(date), "2017-01-01"), 
                     y = cum.precip, 
                     color = factor(year(date)))) +
  geom_path() +
  scale_x_date(date_breaks="months", date_labels="%b") +
  labs(x="",
       colour="",
       y="Cumulative precipitation \n(mm)") +
  theme_bw()

```

## Site conditions 2

```{r}

ggplot(subset(weather, year!="30"), 
                aes(x = as.Date(yday(date), "1970-01-01"), 
                     y = `Cumulative GDD`, 
                     color = factor(year(date)))) +
  geom_line() +
  scale_x_date(date_breaks="months", date_labels="%b") +
  labs(x="",
       colour="",
       y="Cumulative Growing\nDegree Days") +
  theme_bw()


```

## Diagnostics

* missing data
  - grain cuts
  - follow-up cuts @ R70.2018 & I2.2017
* problematic data
  - one outlier value at R100.2019 for dough cut
  - I2 
  
```{r, fig.height=2.5,fig.width=5, fig.align="center"}
lattice::bwplot(yield.1cut~timing.1cut|env,subset(dat,env=="R100.2019" & timing.1cut!="grain"))
```


## I2 removal justification

```{r}
ggplot(subset(dat, timing.1cut!="grain"),
              aes(y=yield.1cut, x=gdd.1cut, color=timing.1cut))+
  facet_grid(~env)+
  geom_point()+
  geom_smooth(method="lm", se=F, show.legend=F, color="black")+
  labs(caption = "n=36 for all environments")
```

## Analysis 1 {.smaller}

**null hypothesis testing**

* repeated measures
  - each plot is measured 1-3 times per year depending on assigned follow.cut treatment
  - each plot is measured over 2 consecutive years
* random effects
  - environment
  - site:block
* fixed effects
  - follow.cut
  - timing.1cut

## Analysis 2 {.smaller}

**regression**

* responses vs.
  - GDD accumulation
  - *precipitation accumulation?*
  - *hydrothermal time?*
  
## Model fitting 1

* Box plot response by environment
    * identify if there's a consistent trend
        * interaction drivers?
    * estimate distribution and variance
* Distribution
    * normal
        - lmer
    * gamma (poisson)
        - glmer
        
## Model fitting 2

* Fit all effects into model
    i) remove random effects not reducing residual variance
    i) compare multiple models
        * log liklihood
        * AIC
    i) null hypothesis test @ alpha=0.05
    i) post-hoc mean comparison
        * tukey, z-statistic


## yield of first cut in second year of data collection

```{r, echo=TRUE}
car::Anova(lmer(yield.1cut~follow.cut*timing.1cut+
                  (1|site:block)+
                  (1|env),
           data=subset(dat1, field.year=="second")))
```
## quality of first cut in second year of data collection 

```{r, echo=TRUE}
car::Anova(lmer(RFQ.1cut~follow.cut*timing.1cut+
                  (1|site:block)+
                  (1|env),
           data=subset(dat1, field.year=="second")))
```


## quality of second cut in second year 

```{r, echo=TRUE}
car::Anova(lmer(RFQ.2cut~follow.cut*timing.1cut+
                  (1|block),
           data=subset(dat1, env=="R100.2019" & follow.cut!="none")))
```

## quality of second cut in both years {.smaller}

```{r, echo=TRUE}
car::Anova(lmer(RFQ.2cut~year*follow.cut+
                  (1|block),
           data=subset(dat1, site=="R100" & follow.cut!="none")))
```
-compare between years at same site since no obvious carryover effect
-conclude that no difference in quality in second cut between the first and second year at R100.

## the effect of multiple cuttings


  dataset                     yield.1cut    rfq.1cut    .2cut     .3cut
----------------------       ------------  ----------  --------  --------
 R70.2018+R100.2019                ns        ns        
 R100.2019                                               ns
 
Table:  Second year datasets do not find differences among follow.cut treatment levels due to the follow.cut treatments applied in the first year.


- no carryover effect from years observed

## follow.cut contribution to yield

```{r}

dat2 <- dat1
dat2$yield.1cut <- dat1$yield.1cut
dat2$yield.2cut <- dat1$yield.1cut+dat1$yield.2cut
dat2$yield.3cut <- dat1$yield.1cut+dat1$yield.2cut+dat1$yield.3cut
dat2.long <- reshape2::melt(data = dat2, id.vars = c("follow.cut"), measure.vars = c("yield.3cut", "yield.2cut","yield.1cut"))

dat2.long <- dat2.long %>%
  mutate(variable = recode(variable,
             yield.3cut = "October cut",
             yield.2cut = "September cut",
             yield.1cut = "Spring cut")) %>%
  mutate(follow.cut = recode(follow.cut,
             october = "september\n+\noctober"))


ggplot(data=dat2.long,
       aes(x=follow.cut, 
           y=value, 
           fill=variable, color=variable)) +
      geom_bar(stat="summary",
               fun="mean",
               position ="identity") +
  labs(y="Cumulative Mean Total Forage Yield \n(kg ha-1 yr-1)",
       x="Follow-Up Cuts",
       fill="",
       color="")

```

-NS differences

## follow.cut contribution to RFQ

```{r}

dat2 <- dat1
dat2$RFQ.1cut <- dat1$RFQ.1cut.wt
dat2$RFQ.2cut <- dat1$RFQ.2cut.wt
dat2$RFQ.3cut <- dat1$RFQ.3cut.wt
dat2.long <- reshape2::melt(data = dat2, id.vars = c("follow.cut"), measure.vars = c("RFQ.3cut", "RFQ.2cut","RFQ.1cut"))

dat2.long <- dat2.long %>%
  mutate(variable = recode(variable,
             RFQ.3cut = "October cut",
             RFQ.2cut = "September cut",
             RFQ.1cut = "Spring cut")) %>%
  mutate(follow.cut = recode(follow.cut,
             october = "september\n+\noctober"))


ggplot(data=dat2.long,
       aes(x=follow.cut, 
           y=value, 
           fill=variable, color=variable)) +
      geom_bar(stat="summary",
               fun="mean",
               position ="stack") +
  labs(y="Average Relative Forage Quality",
       x="Follow-Up Cuts",
       fill="",
       color="")
```

-NS differences

## yield.1cut vs. timing.1cut

```{r}
yield.1cut.tidy <- dat1 %>%
  group_by(timing.1cut) %>%
  summarise(n=n(), mean=mean(yield.1cut), 
            se=sd(yield.1cut)/sqrt(n)) %>%
  as.data.frame() %>%
  mutate(tukey = c("b","a", "a"))

ggplot(yield.1cut.tidy, aes(x=timing.1cut, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.3,
                alpha=1,
                size=1) +
  geom_bar(stat = "identity", width=0.7) +
  geom_text(aes(label=tukey),
            vjust=-1.3,
            size=8) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 5000)) + #adjust for y-axis range 
  theme_classic() +
  labs(y="Forage Yield of the first cut of the Year \n(kg ha-1)",
       x="",
       title="Combined field years")

```

## yield.total vs. timing.1cut

```{r}
dat3 <- subset(dat1,env=="R70.2017" | env=="R100.2018" | env=="R100.2019")

yield.total.tidy <- dat3 %>%
  group_by(treatment) %>%
  summarise(n=n(), mean=mean(yield.total), 
            se=sd(yield.total)/sqrt(n)) %>%
  as.data.frame() %>%
  mutate(tukey = c("b","ab", "ab", "ab", "a", "a", "ab", "ab", "ab"))


ggplot(yield.total.tidy, aes(x=treatment, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.3,
                alpha=1,
                size=1) +
  geom_bar(stat = "identity", width=0.7) +
  geom_text(aes(label=tukey),
            vjust=-1.5,
            size=8) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 10000)) + #adjust for y-axis range 
  theme_classic() +
  labs(y="Forage Yield (kg ha-1)",
       x="",
       title="Combined across three environments")
```

## rfq.1cut vs timing.1cut

```{r}
RFQ.1cut.tidy <- dat1 %>%
  group_by(timing.1cut) %>%
  summarise(n=n(), mean=mean(RFQ.1cut), 
            se=sd(RFQ.1cut)/sqrt(n)) %>%
  as.data.frame() %>%
  mutate(tukey = c("a","b", "c"))

ggplot(RFQ.1cut.tidy, aes(x=timing.1cut, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.3,
                alpha=1,
                size=1) +
  geom_bar(stat = "identity", width=0.7) +
  geom_text(aes(label=tukey),
            vjust=-1.2,
            size=8) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 200)) + #adjust for y-axis range 
  theme_classic() +
  labs(y="Forage Yield (kg ha-1)",
       x="",
       title="")

```
